
```{r setup, echo=FALSE}
library(knitr)
opts_knit$set(out.format = 'html')
opts_chunk$set(highlight = TRUE, tidy = FALSE)
opts_chunk$set(message=FALSE, warning=FALSE, fig.width=5, fig.height=5)
options(scipen=6, digits=4)
``` 
# Class 3 #

---

# Review #

---
* Preparing data in csv format (excel)
* Importing data set from csv.
* Calling variables
* Using **attach()** function
* Telling apart **numeric variables** and **factor variable**
* Using **str()** function
* Using **mean(), median(), sd(), range(), min(), max()** functions
* Storing result in a new variable
* Using **summary()** function
* Getting subset of the data: **subset()** function

---

# 1. More on factors #

---
### Converting to and from factor ###

* Whis is the distinction between numeric and factor variable important?

```{r fig.width=10}
cscore <- read.csv("Class.score.csv")
par(mfrow=c(1,2))
plot(cscore$class, cscore$score)
plot(as.factor(cscore$class), cscore$score)
```

---
```{r}
round(coef(summary(lm(score~class, data=cscore))), 3)
round(coef(summary(lm(score~as.factor(class), data=cscore))), 3)
```
---
* How can we make factors?

```{r}
# We have this character variable names 'ses'
ses <- c("low", "middle", "low", "low", "low", "low", "middle", "low", "middle", "middle", "middle", "middle", "middle", "high", "high", "low", "middle", "middle", "low", "high")

ses.factor<-factor(ses)
ses.factor
```
---
* Why is the factor level "high, low, middle" not "low, middle, high"?

```{r}
# ordered factor
ses.ordered<-ordered(ses, levels=c('low', 'middle', 'high'))
ses.ordered

cscore$class<-ordered(cscore$class, levels=1:10)
cscore$class
```
---
* How can we treat a factor variable as a numeric variable?

```{r}
as.numeric(cscore$class)
```
---

# 2. Correlation #

---
* What is correlation?
: How much of the total variation is explained by co-variation?

```{r}
a<-1:100
b<-a+rnorm(100, 0, sqrt(a))
sample<-data.frame(a,b)
plot(sample)
```

---
```{r}
c<-cov(sample)
c[1,2]/(sqrt(c[1,1])*sqrt(c[2,2]))
cor(sample)
```

---
```{r echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt)

    test <- cor.test(x,y)
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " "))

    text(0.5, 0.5, txt, cex = cex * r)
    text(.8, .8, Signif, cex=cex, col=2)
}
```
* Correlation among multiple variables

```{r}
require(ISwR)
data(cystfibr)
require(ellipse)
plotcorr(cor(cystfibr))
```
---
```{r fig.width=8, fig.height=8}
pairs(cystfibr, lower.panel=panel.smooth, upper.panel=panel.cor)
```
---

# 3. Simple Linear Regression #

---
* Independent variable & Dependent variable
* Explanatory variable or Predictor variable

* How can we **explain** or **predict** dependent variables from **independent variables**?
---
* **Formula notation**
  * independent variable ~ dependent variable(1) + dependent variable(2)....
---
```{r}
plot(sample)
```
---
```{r}
model<-lm(b~a, data=sample)
round(coef(summary(model)), 3)
plot(density(b-mean(b)), ylim=c(0,0.07), main='Density of b and residuals')
lines(density(resid(model)), col='red')
text(50, 0.04, paste('var=', round(var(b-mean(b)),0)))
text(50, 0.035, paste('var=', round(var(resid(model)),0)), col='red')
```